---
title: "functionalizing rasters from NEON"
author: "Kunxuan Wang"
date: "June 21, 2016"
output: html_document
---

# Tasks 
Start by putting up objectives/tasks that students will be working though:
1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

```{r add-libraries}
library(raster)
library(rgdal)
source("scripts/functions.R")

```


```{r import-raster-variable}
# import raster from data
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")
chosen_breaks <- c(6,30,50,100)

```

1. For the CHM, set values == 0 to NA (not trees)

```{r set-NA}
# set 0 values to NA
chm[chm==0] <- NA

```

1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

```{r classify-by-height}

# test_breaks = c(6,30,50,100)
# class_mat = create_h_class_mat(test_breaks)
# class_m_list_test <- c(0, 6, 1,
#                      6, 30, 2,
#                      30, 50, 3,
#                      50,100, 4)
# class_mat_test <- matrix(class_m_list_test, ncol = 3, byrow = TRUE)

#try(if(class_mat != class_mat_test) stop("function create_h_class_mat did not work"))

class_mat = create_h_class_mat(chosen_breaks)

plot_chm_density(chm, "chm density by function", chosen_breaks)

chm_class_h <- reclassify(chm, class_mat)

```



1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

```{r plot-raster}
plot_reclassified_raster(chm_class_h, site.name = "TEAK", chosen_breaks)

```

1. Export the plot figure to a pdf – publishable

```{r export-figure-pdf, eval=FALSE}
make_pdf(plot_reclassified_raster(chm_class_h, site.name = "TEAK", chosen_breaks), 
         filename = "output/TEAK_class_height.pdf", width = 7, height = 6)
```

1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r export-raster}
writeRaster(chm_class_h, 
            filename="TEAK_class_height_chm.tiff",
            format="GTiff",
            NaFlag=-9999,
            overwrite=TRUE)

```

```{r timestamp}
sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"), 
            "_sessionInfo.txt"))
sessionInfo()
sink()
```


